Processing Script

### 2024-02-21 code/eda-code/eda.qmd
Part 2 Updates
I added the skeleton code to produce the remainder of the exploratory plots. I made a note in the processing code about exploring the missing data in the outcome variable.

explore how many people use each method
explore how many countries are in each category for each method using new factor variable
explore percentage of unintended pregnancies across methods

# Data exploration through tables
We will begin with creating a summary table of our cleaned variables using the skim()function. This provides information

```{r}
summary_df = skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

@@

```{r}
p1 <- mydata %>% ggplot(aes(x=pct_upreg)) + geom_histogram() 
plot(p1)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well. We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse`.

@@

We will now explore the new 
plots 6-10

We will start exploring how many women use each type of birth control compared to the percent of unintended pregnancies related to that birth control. Plot 11 shows `pct_upreg_larcster` vs `total_larcster` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.

```{r}
p11 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster)) + geom_point() + geom_smooth(method='lm')
plot(p11)
figure_file = here("results","height_weight.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.

```{r}
p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","height_weight_stratified.png")
ggsave(filename = figure_file, plot=p4) 
```

Our next group of plots will be violin plots to explore the usage level of birth associated with percentage of unintended pregnancies use for each birth control method. 
```{r}
p16 <- mydata %>% ggplot(aes(x=larcster_usage, y=pct_upreg_larcster))
```

Our last plots will be mosaic plots to compare two categorical variables: region and usage for each type of birth control. 
```{r}
library(ggmosaic)
p17 <- mydata %>% ggplot(aes(x=region, y=larcster_usage, fill = freq)) + geom_mosaic(aes(x=region, fill = larcster_usage))
```

>>>>

# Data exploration through tables
We will begin with creating a summary table of our cleaned variables using the skim()function. This provides information about each of the numeric variables including the mean, standard deviation, quartiles, missing observations, and a histogram.

```{r}
summary_df = skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```
We also created simple tables for the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
```{r}

```

@@

```{r}
p1 <- mydata %>% ggplot(aes(x=pct_upreg)) + 
  geom_histogram() + 
  labs(x="Percent of Unintended Pregnancies", y="Frequency at Country Level")
plot(p1)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well. We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.

@@

We will now explore the new total usage variables that were created using the number of women using each method across the four age categories. This set of plots revealed large outliers in the 'total_larcster`, 'total_sarc', 'total_trad', and `total_none` variables, which prevented the graphs from being very useful. The histograms look like most observations for each of these four variables are 0, but that is not possible because the summary table showed that the minimum value for all four of these variables is greater than 0.

probably skewed by population size 
```{r}
p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram(binwidth = 5000) 
plot(p6)
figure_file = here("results", "figures", "exploratory-figures", "total_larcster_distribution.png")
ggsave(filename = figure_file, plot=p6)

p7 <- mydata %>% ggplot(aes(x=total_sarc)) + geom_histogram() 
plot(p7)
figure_file = here("results", "figures", "exploratory-figures", "total_sarc_distribution.png")
ggsave(filename = figure_file, plot=p7)

p8 <- mydata %>% ggplot(aes(x=total_trad)) + geom_histogram() 
plot(p8)
figure_file = here("results", "figures", "exploratory-figures", "total_trad_distribution.png")
ggsave(filename = figure_file, plot=p8)

p9 <- mydata %>% ggplot(aes(x=total_none)) + geom_histogram() 
plot(p9)
figure_file = here("results", "figures", "exploratory-figures", "total_none_distribution.png")
ggsave(filename = figure_file, plot=p9) 
```

We will start exploring how many women use each type of birth control compared to the percent of unintended pregnancies related to that birth control. Plot 10 shows `pct_upreg_larcster` vs `total_larcster` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods. Plot 11 uses `pct_upreg_sarc` and `total_sarc` to examine the same relationship as plot 10 but with short-acting reversible methods. 
Plot 12 shows `pct_upreg_trad` versus `total_trad`, and plot 13 shows `pct_upreg_nouse` versus `total_none`. 
```{r}
p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster)) + geom_point() + geom_smooth(method='lm')
plot(p10)
figure_file = here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png")
ggsave(filename = figure_file, plot=p10)

p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=total_sarc)) + geom_point() + geom_smooth(method='lm')
plot(p11)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=p11)

p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=total_trad)) + geom_point() + geom_smooth(method='lm')
plot(p12)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png")
ggsave(filename = figure_file, plot=p12)

p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=total_none)) + geom_point() + geom_smooth(method='lm')
plot(p13)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png")
ggsave(filename = figure_file, plot=p13) 
```

We are also interested in how these relationships between percent of unintended pregnancies using each birth control method and the number of people using each method varies across region. We adjusted the original scatter plots created above by coloring based on region.

```{r}
p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster, color = region)) + geom_point() + geom_smooth(method='lm')
plot(p14)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png")
ggsave(filename = figure_file, plot=p14)

p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=total_sarc, color = region)) + geom_point() + geom_smooth(method='lm')
plot(p15)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=p15)

p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=total_trad, color = region)) + geom_point() + geom_smooth(method='lm')
plot(p16)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png")
ggsave(filename = figure_file, plot=p16)

p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=total_none, color = region)) + geom_point() + geom_smooth(method='lm')
plot(p17)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png")
ggsave(filename = figure_file, plot=p17) 
```

Our next group of plots will be violin plots to explore the usage level of each birth control message associated with percentage of unintended pregnancies use for each birth control method. 
```{r}
p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
  geom_violin() +
  labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
plot(p18)
figure_file = here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p18)

p19 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
  geom_violin() +
  labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
plot(p19)
figure_file = here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p19)

p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
  geom_violin() +
  labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
plot(p20)
figure_file = here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p20)

p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
  geom_violin() +
  labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
plot(p21)
figure_file = here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p21) 
```

### 2024-02-21 code/processing-code/processingfile-v1.qmd
Part 2 Updates
I added the skeleton code to produce the remainder of the exploratory plots. I made a note in the processing code about exploring the missing data in the outcome variable.

We should impute the missing data here to avoid introducing bias because 18.18% of the data is missing, and this variable is essential for answering the research question. After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.  
```{r}
#| message: false
# find mean to use for imputed data
summary(rawdata2$pct_upreg)

>>>>

We should impute the missing data here to avoid introducing bias because 18.18% of the data is missing, and this variable is essential for answering the research question. After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data. `mice` package or `tidymodels`
or remove because it is main outcome
or make our own pct_upreg_updated and compare to what they have
```{r}
#| message: false
# find mean to use for imputed data
summary(rawdata2$pct_upreg)
